Tagging variables with keywords

### 2020-07-17 data-raw/variable-keywords-table.Rmd
just one space
one comment
update to variable-keywords-table with zacat variable groups from EB/44.2bis(1996) and other fixes

```{r setup}
library(eurobarometer)
library(tidyverse)
library(tibble)
library(knitr)
library(kableExtra)
library(tidytext)
library(janitor)
library(stringi)
```

@@

@@

Mapping of GESIS Archive IDs (as used in file names) to wave numbers, from EB/44.2bis (ZA2828) to EB/92.1 (ZA7579) - 156 EB waves altogether.

```{r, warning = F}
eb_names <- rio::import("../data-raw/eb_waves_filenames.xlsx") %>%
  separate(x, c("archive_id", "x2", "title"), sep = ":") %>%
  separate(x2, c("wave", "dates"), sep = " \\(|\\)") %>%
  mutate(wave = gsub("Eurobarometer | ", "", wave))

@@

All procedures below applied to 51 waves, for which all metadata are available.

@@

```{r}
dict <- rio::import("../data-raw/var_lab_dict.xlsx", sheet = "dict") %>%
  gather(id, from, 2:6) %>%
  select(from, to) %>%
  drop_na(from)

@@

dat2 <- metadata_database %>%
  # filter only waves for which variable groups are available
  filter(archive_id %in% archive_ids,
         # exclude this strange variable
         var_label_orig != "filename") %>%
  # keep one row per variable
  distinct(filename, qb, var_name_orig, var_name_orig, var_label_orig, var_label_norm, archive_id) %>%
  # original variable labels in the data and in GESIS ZACAT are not exactly the same 
  # --> need some cleaning
  mutate(var_label_orig2 = gsub("^W[0-9]{1,2}", "", var_label_orig),
         var_label_orig2 = gsub("^W[0-9]{1,2}[A-Z]", "", var_label_orig),
         var_label_orig2 = gsub("^[0-9]{1-3} ", "", var_label_orig2),
         var_label_orig2 = gsub("^[A-Z][0-9]{1,2}[A-Z] ", "", var_label_orig2),
         var_label_orig2 = gsub("^[A-Z]{1,2}[0-9]{1,2} ", "", var_label_orig2),
         var_label_orig2 = gsub("^[A-Z]{1,2}[0-9]{1,2}[A-Z] ", "", var_label_orig2),
         var_label_orig2 = gsub("^[0-9]{1,2} ", "", var_label_orig2),
         var_label_orig2 = gsub("^[A-Z][0-9] [0-9]{1,2} ", "", var_label_orig2),
         var_label_orig2 = gsub("^W3A/W4A ", "", var_label_orig2)) %>%
  # join with table with variable groups
  left_join(varlabs, by = c("var_label_orig2" = "var_label_zacat", "archive_id")) %>%
  # remove variables in the following groups
  filter(!group %in% c("Archive and survey ID variables",
                       "Country group variables",
                       "Nationality (Q1)",
                       "Protocol variables (P)",
                       "Standard nation ID variables",
                       "Weight variables (W)",
                       "Nation ID variables",
                       "Protocol variables",
                       "Weight variables",
                       "Archive and ID variables",
                       "Household Weights"),
         # remove recoded variables
         !grepl("recoded|(rec)", var_label_norm)) %>%
  mutate(var_id = paste0(filename, " ", var_name_orig),
         # leftover question numbers at the beginning of value labels
         var_label_norm1 = gsub("^w[0-9]{1,2}", "", var_label_norm),
         var_label_norm1 = gsub("^[0-9]{1-3}_", "", var_label_norm1),
         var_label_norm1 = gsub("^[a-z][0-9]{2}[a-z]_", "", var_label_norm1),
         var_label_norm1 = gsub("^[a-z]{1,2}[0-9]{1,2}_", "", var_label_norm1),
         var_label_norm1 = gsub("^[a-z]{1,2}[0-9]{1,2}[a-z]_", "", var_label_norm1),
         var_label_norm1 = gsub("^[0-9]{1,2}_", "", var_label_norm1),
         var_label_norm1 = gsub("^[a-z]_", "", var_label_norm1),
         var_label_norm1 = gsub("^[a-z][0-9]_[0-9]{1,2}_", "", var_label_norm1)) %>%
  # apply the translation table
  mutate(var_label_norm1 = stri_replace_all_regex(var_label_norm1,
                                                  pattern = dict$from,
                                                  replacement = dict$to,
                                                  vectorize_all = FALSE)) %>%
  # replace underscore with space
  mutate(text = gsub("_", " ", var_label_norm1)) %>%
  ungroup() %>%
  # select the needed variables
  select(filename, qb, var_label_orig, var_label_norm, var_label_norm1, text, var_id, group)

@@

tidy <- dat2 %>%
  # tokenize
  unnest_tokens(word, text, 
                token = "regex", pattern = " |  ") %>%
  # eliminate stop words, with some exceptions
  anti_join(stop_words %>%
              filter(!word %in% c("working", "right", "other")
                     )) %>%
  # eliminate numbers
  filter(!grepl("\\b\\d+\\b", word)) %>%
  # convert country codes to country names
  mutate(
    country = ifelse(word %in% eu_countries, word, NA),
         country = countrycode::countrycode(
           toupper(country), "iso2c", "country.name"),
    country = tolower(country),
    word = ifelse(!is.na(country), country, word),
    word = ifelse(word == "united kingdom", "united-kingdom", word),
    word = ifelse(word == "tcc", "turkish-cypriot-community", word)
    ) %>%
  select(-country)
```

@@

Filter out labels that have words that appear in < 5 different waves. If we want to create a trend file, we're only interested in variables that are repeated often.

```{r}
tidy_small_2 <- tidy %>%
  group_by(word) %>%
  # number of distinct file names (waves) per word
  mutate(n_wave = n_distinct(filename)) %>%
  # flag words that appear in < 5 different files
  mutate(to_drop = as.numeric(n_wave < 10)) %>%
  # group by variable ID (file x variable name)
  group_by(var_id) %>%
  # copy to_drop across words in the same variable label
  mutate(to_drop = max(to_drop, na.rm = TRUE)) %>%
  # drop variable labels with the to_drop identifier
  filter(to_drop != 1) %>%
  # remove auxiliary variables
  select(-to_drop, -n_wave) %>%
  ungroup()

@@

```{r}
# get the order of variables according to declining proportion of missing value
# (declining popularity of keywords)
order <- dtm %>%
  summarise_all(funs(sum(is.na(.))/length(.))) %>%
  t() %>% as.vector()

@@

```{r}
dtm2 <- dtm %>%
  select( names(sort(order)) ) %>%
  mutate_at(vars(3:ncol(dtm)), 
            funs(ifelse(is.na(.), "", .))
            ) %>%
  unite("var_label_std", 3:ncol(dtm)) %>%
  mutate(var_label_std = gsub(
    "(\\_)\\1+", "\\1", var_label_std) ) %>%
  mutate(var_label_std = gsub(
    "^_|_$", "", var_label_std) )

@@

These standardized variable names still have problems that need to be fixed.

@@

```{r}
# filter on var_label_std
dtm2 %>%
  filter(grepl("trust|satisf|left-right|interest|class", var_label_std)) %>% 
  count(var_label_std) %>%
  arrange(desc(n)) %>%
  head(., 25) %>%
  kable %>%
  kable_styling(bootstrap_options =
                  c("striped", "hover", "condensed"),
                  fixed_thead = T,
                  font_size = 10 )
```

>>>>

```{r setup}
library(eurobarometer)
library(rio)
library(tidyverse)
library(tibble)
library(knitr)
library(kableExtra)
library(tidytext)
library(janitor)
library(stringi)
```

@@

These 97 include 96 Standard Eurobarometer waves and 1 Candidate Country EB wave.

@@

Mapping of GESIS Archive IDs (as used in file names) to wave numbers, from EB/44.2bis (ZA2828) to EB/92.1 (ZA7579) - 157 EB waves altogether.

```{r, warning = F}
eb_names <- import( file.path('../data-raw', 'eb_waves_filenames.xlsx')) %>%
  separate(x, c("archive_id", "x2", "title"), sep = ":") %>%
  separate(x2, c("wave", "dates"), sep = " \\(|\\)") %>%
  mutate(wave = gsub("Eurobarometer | ", "", wave))

@@

All procedures below applied to those waves, for which all metadata are available.

@@

```{r}
dict <- import( file.path('../data-raw', 'var_lab_dict.xlsx'), 
                sheet = "dict") %>% 
  gather(id, from, 2:6) %>%
  select(from, to) %>%
  drop_na(from)

@@

dat2 <- metadata_database %>%
  # filter only waves for which variable groups are available
  filter(archive_id %in% archive_ids,
         # exclude this strange variable
         var_label_orig != "filename") %>%
  # keep one row per variable
  distinct(filename, qb, var_name_orig, var_name_orig, var_label_orig, var_label_norm, archive_id) %>%
  # original variable labels in the data and in GESIS ZACAT are not exactly the same 
  # --> need some cleaning
  mutate(var_label_orig2 = gsub("^([A-Z])*[0-9]{1,3}([A-Z]{1,2})* ", "", 
                                var_label_orig),
         var_label_orig2 = gsub("^[A-Z][0-9] [0-9]{1,2} ", "", 
                                var_label_orig2),
         var_label_orig2 = gsub("^W3A/W4A ", "", var_label_orig2),
         var_label_orig2 = gsub("^[A-Z]{1,2}[0-9]{1,2}([A-Z])*_[A-Z]{1,2} ", "", 
                                var_label_orig2),
         var_label_orig2 = gsub("^\\s+|\\s+$", "", 
                                var_label_orig2),
         var_label_orig2 = gsub('“)', "", 
                                var_label_orig2),
         var_label_orig2 = gsub("  ", " ", 
                                var_label_orig2),
         # these 2 waves have problems and labels in data and in ZACAT are different
         var_label_orig2 = ifelse(archive_id %in% c("ZA6695", "ZA6697"),
                                  gsub("\\s*\\([^\\)]+\\)","", var_label_orig2),
                                  var_label_orig2)) %>%
  # join with table with variable groups
  left_join(varlabs, by = c("var_label_orig2" = "var_label_zacat", "archive_id")) %>%
  # remove variables in the following groups
  filter(!grepl("protocol|weight|nationality|nation id|original id|id variable|country group", 
                tolower(group)),
         # remove recoded variables
         !grepl("recoded|(rec)", var_label_norm)) %>%
  mutate(var_id = paste0(filename, " ", var_name_orig),
         # leftover question numbers at the beginning of value labels
         var_label_norm1 = gsub("^w[0-9]{1,2}", "", 
                                var_label_norm),
         var_label_norm1 = gsub("^[0-9]{1-3}_", "", 
                                var_label_norm1),
         var_label_norm1 = gsub("^[a-z][0-9]{2}[a-z]_", "", 
                                var_label_norm1),
         var_label_norm1 = gsub("^[a-z]{1,2}[0-9]{1,2}_", "", 
                                var_label_norm1),
         var_label_norm1 = gsub("^[a-z]{1,2}[0-9]{1,2}[a-z]_", "", 
                                var_label_norm1),
         var_label_norm1 = gsub("^[0-9]{1,2}_", "", 
                                var_label_norm1),
         var_label_norm1 = gsub("^[a-z]_", "", 
                                var_label_norm1),
         var_label_norm1 = gsub("^[a-z][0-9]_[0-9]{1,2}_", "", 
                                var_label_norm1)) %>%
  # apply the translation table
  mutate(var_label_norm1 = stri_replace_all_regex(var_label_norm1,
                                                  pattern = dict$from,
                                                  replacement = dict$to,
                                                  vectorize_all = FALSE)) %>%
  ungroup() %>%
  # select the needed variables
  select(filename, qb, var_label_orig, var_label_orig2, 
         var_label_norm, var_label_norm1, var_id, group)

@@

# ---- but normally one should be in attr(x, "na_values")

tidy <- dat2 %>%
  # tokenize
  unnest_tokens(word, var_label_norm1, 
                token = "regex", 
                pattern = "_",
                drop = FALSE) %>%
  # eliminate stop words, with some exceptions
  anti_join(stop_words %>%
              filter( !word %in% c("working", "right", "other")
                     )) %>%
  # eliminate numbers
  filter( !grepl("\\b\\d+\\b", word)) %>%
  # convert country codes to country names
  mutate(
    country = ifelse(word %in% eu_countries, 
                     word, 
                     NA),
         country = countrycode::countrycode(
           toupper(country), "iso2c", "country.name"),
    country = tolower(country),
    word = ifelse(!is.na(country), 
                  country, 
                  word),
    word = ifelse(word == "united kingdom", 
                  "united-kingdom", 
                  word),
    word = ifelse(word == "tcc", 
                  "turkish-cypriot-community", 
                  word)
    ) %>%
  select(-country)
```

@@

Filter out labels that have words that appear in < 10 different waves. If we want to create a trend file, we're only interested in variables that are repeated often.

```{r}
tidy_small_2 <- tidy %>%
  group_by(word) %>%
  # number of distinct file names (waves) per word
  mutate(n_wave = n_distinct(filename)) %>%
  # flag words that appear in < 10 different files
  mutate(to_drop = as.numeric(n_wave < 10)) %>%
  # group by variable ID (file x variable name)
  group_by(var_id) %>%
  # copy to_drop across words in the same variable label
  mutate(to_drop = max(to_drop, na.rm = TRUE)) %>%
  # drop variable labels with the to_drop identifier
  filter(to_drop != 1) %>%
  # remove auxiliary variables
  select(-to_drop, -n_wave) %>%
  ungroup()

@@

```{r}
# get the order of variables according to declining proportion of missing value
# (declining popularity of keywords)
order <- dtm %>%
  summarise_all(
    funs(sum(is.na(.))/length(.))
    ) %>%
  t() %>% as.vector()

@@

!!! This is not a perfect solution for the question order. We need a table with "concepts", like trust, satisfaction, etc., so that they are always at the beginning. Otherwise we end up with "internet_trust" (which should be "trust_internet") and so on.

```{r}
dtm2 <- dtm %>%
  select( names(sort(order)) ) %>%
  mutate_at( vars(3:ncol(dtm) ), 
            funs(ifelse(is.na(.), "", .))
            ) %>%
  unite("var_label_std", 3:ncol(dtm)) %>%
  mutate(var_label_std = gsub(
    "(\\_)\\1+", "\\1", var_label_std) ) %>%
  mutate(var_label_std = gsub(
    "^_|_$", "", var_label_std) )

@@

!!! These standardized variable names still have problems that need to be fixed.

@@

```{r}
# filter on var_label_std
dtm2 %>%
  filter( grepl("trust|satisf|left-right|interest|class",
                var_label_std)) %>% 
  count(var_label_std) %>%
  arrange(desc(n)) %>%
  head(., 25) %>%
  kable %>%
  kable_styling(bootstrap_options =
                  c("striped", "hover", "condensed"),
                  fixed_thead = T,
                  font_size = 10 )
```
